### 2018-01-03 Tex/presentations/SPSA2018/DesmaraisSPSATalk.tex
adding SPSA slides

\documentclass[handout]{beamer}
\usepackage{bm}
\usepackage{listings}
\usepackage{algorithmic}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{multirow}
\lstloadlanguages{R}
\lstset{ language=R, basicstyle=\scriptsize\ttfamily, commentstyle=\ttfamily\color{gray}, numbers=left, numberstyle=\ttfamily\color{gray}\footnotesize, stepnumber=1, numbersep=5pt, backgroundcolor=\color{white}, showspaces=false, showstringspaces=false, showtabs=false, frame=single, tabsize=2, captionpos=b, breaklines=true, breakatwhitespace=false, escapeinside={}, keywordstyle={}, morekeywords={} }\renewcommand{\P}{\mathcal{P}}
\newcommand{\bt}{\pmb{\theta}}
\newcommand{\bG}{\pmb{\Gamma}}
\newcommand{\p}{\pause}
\newcommand{\pitem}{\pause \item}
\newcommand{\N}{\mathcal{N}}
\newcommand{\Y}{\bm{\mathcal{Y}}}
\newcommand{\bh}{\bm{h}} 
\DeclareMathOperator*{\argmax}{arg\,max}

\usefonttheme{serif}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

%
\title{Inferential Analysis of the  \\ Supreme Court Citation Network}
%
\definecolor{umassred}{HTML}{001769}
\setbeamercolor{structure}{fg=umassred} 
\author{ \textbf{Bruce A. Desmarais}\\ Collaborators: Christian Schmid, Ted Chen, David Hunter \\~\\ {\bf Pennsylvania State University}}

%\affil[1]{University of Massachusetts Amherst}
%\affil[2]{University of Connecticut}

\begin{document}
%\date{November 16, 2012}

\begin{frame}
\vspace{1cm}
  \titlepage
  \vspace{0cm}
  \begin{center}
   \begin{tabular}{cc}
\hspace*{-.2in} \tiny \begin{minipage}{3.5in}
Work supported by NSF grants SES-1558661, SES-1619644, SES-1637089, and CISE-1320219)\\ ~\\~\\~\\~\\
\end{minipage}
& \includegraphics[scale=.05]{figures/NSF_logo.png}
\end{tabular}
\end{center}
\end{frame}

\begin{frame} \frametitle{Captivating Story/Anecdote}

\end{frame}

\begin{frame} \frametitle{Two Limitations in Case-Level Analysis}

\end{frame}

\begin{frame} \frametitle{Reciprocity}

\end{frame}

\begin{frame} \frametitle{Transitivity}

\end{frame}

\begin{frame} \frametitle{Popularity and Sociality}

\end{frame}

\begin{frame} \frametitle{Citation TERGM}

\end{frame}

\begin{frame} \frametitle{Model Specification}

\end{frame}

\begin{frame} \frametitle{Results Dependence}

\end{frame}

\begin{frame} \frametitle{Results Covariates}

\end{frame}

\begin{frame} \frametitle{Predictive Performance}

\end{frame}

\begin{frame} \frametitle{Conclusion}

\end{frame}

\end{document}